---
title: "cleaning wildlife insights data"
output: html_document
date: "2023-02-19"
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
library(lubridate)
library(janitor)
library(here)
```


Data Cleaning TO do:
in common name column
- replace chadriiformes + calidri species with shorebird
- replace Neotoma species with sp

remove all data with dates earlier than 2022
UPDATE: fixed all dates in wildlife insights as of 23 feb 23

UGH WI HAS MID-ID'D SOME DATES


```{r data import}

wild <- read_csv(here("data/WI_images23feb23.csv")) %>% 
  select(!c(project_id, image_id, location, wi_taxon_id, uncertainty, age, sex, animal_recognizable, individual_id, individual_animal_notes, highlighted, cv_confidence, license)) %>% 
  mutate(timestamp = ymd_hms(timestamp)) %>% 
  separate(timestamp, into = c('date', 'time'), sep=' ', remove = TRUE) %>% 
  mutate(year = year(date)) %>% 
  mutate(month = month(date, label = TRUE)) %>% 
  uncount(number_of_objects, # "uncounts" i.e. makes specified number of copies of the row based on the # in how_many
          # we are basically multipling the number seconds of animal activity by the number of animals in the capture - this is fine...
          
          .remove = FALSE #check your work by including .remove = FALSE... this keeps the column where the #of copies to make was stored. default is to delete it
          ) %>% 
  filter(month != "Dec") %>% 
  filter(year <= 2022) # eventually can take this out 

unique(wild$deployment_id)

```

Summarize the data for some quick visualizations
```{r making summary tables}

species_site <- wild %>%
  count(deployment_id, common_name) # makes a summary table of # animal-seconds per species per deployment_id per site
########################################################################################

######### make table of total animal activity (seconds = incidences) by deployment_id ###########
count_site <- wild %>%
  count(deployment_id)
########################################################################################

######### make table of animal activity (seconds = incidences) by species summed across all deployment_ids ###########
count_species <- wild %>%
  count(common_name) # makes a summary table of # animal-seconds per species per deployment_id per month
########################################################################################


######### make table of animal activity (seconds = incidences) by deployment_id by date ###########
count_site_date <- wild %>%
  count(deployment_id, month) # makes a summary table of # animal-seconds per species per deployment_id per month
########################################################################################


######### make table of animal activity (seconds = incidences) summed across all deployment_ids by date ###########
count_month <- wild %>%
  count(month) # makes a summary table of # animal-seconds per month
########################################################################################

```

##Now that data is clean, we can visualize. 
###First, animal sightings by site 
###Then, animal sightings by species
###Finally, make a map of pie charts breaking down species sightings in space
```{r number of sightings by site of all animals}

# set order of sites by distance from Jalama
count_site$deployment_id <- factor(count_site$deployment_id, levels = c("Boathouse Cam", "Sudden Canyon Cam", "Old Fencepost Cam" , "Black Canyon Cam", "North Beach Fort Cam", "North Vista Spring Cam", "North Beach Canyon Cam","Cove Cam", "Govies Cliff Cam", "Big Cojo Cam"  ,"Percos Driftwood Cam" , "Percos Boat Cam","Percos Beach Cam", "Percos Log Cam" , "Percos Post Cam"  , "Damsite Creek Cam")) # arranged them north to south
 

countbysite <- ggplot(count_site, aes(deployment_id, n)) +
  geom_bar(stat = "identity") + # so i can pass a value to the bar instead of having ggplot count for me
  theme_classic()+
   theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 10),
        axis.text.y = element_text(size = 20),
        axis.title.y = element_text(size = 20),
        axis.title.x = element_blank())+
  ylab("Number of Individual Sightings")+
  xlab("Site - North to South")

```

```{r number of sightings by species}

countbyspecies <- ggplot(count_species, aes(x = reorder(common_name, -n), n)) +
  geom_bar(stat = "identity") + # so i can pass a value to the bar instead of having ggplot count for me
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 7),
        axis.text.y = element_text(size = 30),
        axis.title.y = element_text(size = 30),
        axis.title.x = element_blank())+
  ylab("Number of Sightings")+
  xlab("Species")

```
```{r number of sightings by month}

countbymonth <- ggplot(count_month, aes(x = month, y = n)) +
  geom_bar(stat = "identity") + # so i can pass a value to the bar instead of having ggplot count for me
  theme_classic()+
  theme(axis.text.x = element_text(angle = 45, 
                                   hjust = 1, 
                                   size = 7),
        axis.text.y = element_text(size = 30),
        axis.title.y = element_text(size = 30),
        axis.title.x = element_blank())+
  ylab("Number of Sightings")+
  xlab("Month")

```

```{r stacked bar chart of species by site}

# set order of sites by distance from Jalama
species_site$deployment_id <- factor(species_site$deployment_id, levels = c("Boathouse Cam", "Sudden Canyon Cam", "Old Fencepost Cam" , "Black Canyon Cam", "North Beach Fort Cam", "North Vista Spring Cam", "North Beach Canyon Cam","Cove Cam", "Govies Cliff Cam", "Big Cojo Cam"  ,"Percos Driftwood Cam" , "Percos Boat Cam","Percos Beach Cam", "Percos Log Cam" , "Percos Post Cam"  , "Damsite Creek Cam")) # arranged them north to south

stacked_spp_site <- ggplot(species_site, aes(deployment_id, n)) +
  geom_bar(aes(fill = common_name), # lets me make clustered bar chart
           stat = "identity", # idk but necessary
           #position = "dodge" # otherwise default is stacked barchart
           )+
  theme_classic()+ 
  ylab("Number of Sightings")+
  xlab("Sites North to South")+
  theme(axis.text.x = element_blank(),
    #axis.text.x = element_text(),
        axis.text.y = element_text(size = 30),
        axis.title.y = element_text(size = 30),
        axis.title.x = element_text(size = 30))

```

### We can also visualize human data by sourcing some previously written code
```{r human data}
source("DataWrangling/human_count_script.R") # a lil script i wrote to get the human data into a nice neat little summary table of human seconds by site. it's called hooman_count

human_clean <- hooman_count %>% 
  #filter(date == "10may22" | date == "14may22" | date == "17apr22" | date == "28apr22" | date == "5apr22") %>% 
  # we only wnt human counts that match the current timeline we're working with, may and april
  mutate(site = case_when(
    site == "perb" ~ "Percos Beach Cam",
    site == "gc" ~ "Government Point Cam",
    site == "gc2" ~ "Government Point Cam",
    site == "bc" ~ "Black Canyon Cam",
    site == "big" ~ "Big Cojo Cam",
    site == "boat" ~ "Percos Boat Cam",
    site == "boatcam17apr22" ~ "Percos Boat Cam",
    site == "bone" ~ "Boneyard Cam",
    
    TRUE ~ site
  )) %>% 
  filter(!is.na(site)) %>% 
  group_by(site) %>% 
  summarise(human = sum(n)) #gets total count of humans by site

######################## merge human and animal data into one df ########################

humans_animals_bysite <- left_join(count_site, human_clean, all = TRUE) %>% 
  replace_na(list(human = 0)) %>% 
  pivot_longer(cols = n:human, names_to = "human_or_animal", values_to = "count") # pivot for ggplot (im crying)

humans_animals_site$site <- factor(humans_animals_site$site, levels = c("vsc", "vof", "bc", "fort", "nvs", "nbc","cov", "gov", "pd", "pb", "pl", "pp", "dam")) # arranged them north to south

```

## Want to grab some numbers on coyotes only

```{r coyote stuff}

coyote <- wild %>% 
  filter(common_name == "Coyote") %>% 
  distinct(date, time, .keep_all = TRUE) # only keep unique combos of date and time to reduce the 8x thing

coyote_sum <- coyote %>% 
  count()

```

