---
title: "Manifest"
author: "Zoe"
date: "5/2/2022"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)

library(tidyverse)
```

#Making a manifest
## this is necessary for importing files to Zooniverse
## the goal is to get a dataframe with every single filename of every photo we are importing
## subjects are rows, each subject has 8 photos

## For my own edification, I want to also add a column that denotes location, site name, and date

### the file path for the folder where the images currently are CANT stay the same -- /Volumes/GoogleDrive/My Drive/SCHOOL/PHD_AT_UCSB/GaviotaCameraTrapping/Zooniverse/Zooniverse_Metadata/....

### the code below only applies to Spring2022. future code will be simpler because I'll try to be less of an idiot

```{r}
# we will use list.files() to find the filenames

filenames <- list.files("/Volumes/GoogleDrive/My Drive/SCHOOL/PHD_AT_UCSB/GaviotaCameraTrapping/Zooniverse/animal_images_8/Spring2022",
                        include.dirs = FALSE,
                        full.names = FALSE,
                        all.files = FALSE)


manifest <- read.table(text = filenames)  %>% # turns filename char vec into dataframe
  distinct() %>%    # removes dups which happen when drive creates temp files
  
  ## need to fix the fact that filenames are in two different formats because I'm an idiot and it's hard to batch fix shit like this outside of command line:
  mutate(V2 = str_replace_all(V1, c("bc17apr22" = "bc_17apr22",
                                    "bc28apr22" = "bc_28apr22",
                                    "cove17apr22" = "cov_17apr22",
                                    "gov17apr22" = "gov_17apr22", 
                                    "nvs28apr22" = "nvs_28apr22",
                                    "pb28apr22" = "pb_28apr22",
                                    "pl17apr22" = "pl_17apr22",
                                    "pp17apr22" = "pp_17apr22",
                                    "gc28apr22" = "gc_28apr22",
                                    "pd28apr22" = "pd_28apr22")
                              )
         ) %>%  # not sure why this is the syntax but for str_replace for a dataframe this is what works?
  
  separate(V2, c("site", "date", "photonum"), sep = "_", remove = FALSE) %>%  # separates filename into site/date identifier and orig photo number
  mutate(photonum = str_remove(photonum, ".JPG"))  # remove the the .JPG from photonum
  mutate(photonum = as.numeric(photonum)) %>% 
  arrange(sitedate, photonum) %>% 
  mutate(imgnum = rep_len(1:8, nrow(.))) %>%  # adds a column repeating 1:8 # CAROLINE SAYS YOU NEED THE DOT I HATE HER
  mutate(subject = rep(1:(nrow(.)/8), each = 8)) %>%   # adds a column linking all images related to one subj together
### now need to expand the dataframe long ways so that 1:8 are 8 different columns
  # need to reorder the dataframe by sitedate and THEN photonum
  select(!photonum) %>% # have to take photonum out because it adds a unique identifier to each photo which messes up pivot_wider
  pivot_wider(names_from = imgnum, values_from = V1, names_prefix = "image") # creates a new column (image1-image8) for each of the 8 images per subject (V1 is the col of filenames)
            





  
write_csv(manifest,"/Volumes/GoogleDrive/My Drive/SCHOOL/PHD_AT_UCSB/GaviotaCameraTrapping/Zooniverse/animal_images/manifest.csv") # now we have a table of all the filenames as a csv
```

